---
title: "Project_SingRs"
output: html_document
---

##Acknowledgement: Frederico Belo, Ahmed Guecioueur

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages
```{r, echo=FALSE}
if("pacman" %in% rownames(installed.packages()) == FALSE) {install.packages("pacman")} # Check if you have universal installer package, install if not

pacman::p_load("caret","ROCR","lift","glmnet","MASS","e1071",'dplyr',
               "partykit","rpart", "munsell","caTools", "caret", "randomForest") #Check, and if needed install the necessary packages
```

## Prediction of stock movements on stock exchange 

We will do the analysis of past share prices of the S&P500 companies, trading volumes, bid low, ask high, price volatiilty to predict the movement of the stock next day. 

- Data would be used from WRDS CRSP

```{r data_loading}
data_general <- read.csv("oracle_data_3.csv")
data_working <- data_general
str(data_working)
```


##Clean the data
Bidlo
 - check for 0s (no price available)
 
```{r, check Bidlo}
which(c(data_working$BIDLO.T.5,data_working$BIDLO.T.4, data_working$BIDLO.T.3, data_working$BIDLO.T.2, data_working$BIDLO.T.1, data_working$BIDLO.T.0) == 0, arr.ind = TRUE)
# confirm no 0s
```

Askhi
 - check for 0s (no price available)
```{r, check Askhi}
which(c(data_working$ASKHI.T.5,data_working$ASKHI.T.4, data_working$ASKHI.T.3, data_working$ASKHI.T.2, data_working$ASKHI.T.1, data_working$ASKHI.T.0) == 0, arr.ind = TRUE)
# Confirm no 0s
```

Numtrd 
 - check for 99 (means number unavailable)
```{r, check Numtrd}
which(c(data_working$NUMTRD.T.5, data_working$NUMTRD.T.4, data_working$NUMTRD.T.3, data_working$NUMTRD.T.2, data_working$NUMTRD.T.1, data_working$NUMTRD.T.0) == 99, arr.ind = TRUE)
# Confirm no 99s
```
 
##Feature engineerring
Create log returns + lag
Create volatility

###Split the data into testing and training
```{r data_split_cleaning}
data_working$date <- NULL

#data_working$date <- as.Date(data_working$date, "%d.%b.%y")
#data_working$VOL.T.10 <- as.numeric(data_working$VOL.T.10)
#data_working$VOL.T.9 <- as.numeric(data_working$VOL.T.9)
#data_working$VOL.T.8 <- as.numeric(data_working$VOL.T.8)
#data_working$VOL.T.7 <- as.numeric(data_working$VOL.T.7)
#data_working$VOL.T.6 <- as.numeric(data_working$VOL.T.6)
#data_working$VOL.T.5 <- as.numeric(data_working$VOL.T.5)
#data_working$VOL.T.4 <- as.numeric(data_working$VOL.T.4)
#data_working$VOL.T.3 <- as.numeric(data_working$VOL.T.3)
#data_working$VOL.T.2 <- as.numeric(data_working$VOL.T.2)
#data_working$VOL.T.1 <- as.numeric(data_working$VOL.T.1)
data_working$MOVEMENT.T.0 <- as.factor(data_working$MOVEMENT.T.0)

#data_working<-fixNAs(data_working)
#data_working<-fixNegatives(data_working)


#data_working_wT0 <- data_working[,1:50]
#data_working_wT0 <- cbind(data_working_wT0, data_working$MOVEMENT.T.0)
#names(data_working_wT0)[51] <- paste("MOVEMENT.T.0")

train_general <- data_working[1:2000,]
test_general <- data_working[2001:2764,]
test_general_womov <- test_general
test_general_womov$MOVEMENT.T.0 <- NULL
```


```{r data_split_2}

set.seed(77850) #set a random number generation seed to ensure that the split is the same everytime
inTrain <- createDataPartition(y = train_general$MOVEMENT.T.0,
                               p = 0.75, list = FALSE)
training <- train_general[ inTrain,]
testing <- train_general[ -inTrain,]
```


###Data engenieering

Transform dataset and take share price for the past 30days
Vaiarbles would be volume traded, return, volatility each day. Shareprice would be independent variable
rows - would be timeseries

###Random forest


```{r random_forest}
start_time <- Sys.time()
model_forest <- randomForest(MOVEMENT.T.0~ ., data=training, importance=TRUE, proximity=TRUE, type="classification")
print(model_forest)
end_time <- Sys.time()
end_time - start_time

plot(model_forest)
importance(model_forest)
varImpPlot(model_forest)
```


Finding predicitons: probabilities and classification

```{r rf_predictions}
forest_probabilities<-predict(model_forest,newdata=testing,type="prob") 
forest_classification<-rep("0",length(testing$MOVEMENT.T.0))
forest_classification[forest_probabilities[,2]>0.2]="1"
forest_classification<-as.factor(forest_classification)
confusionMatrix(forest_classification,testing$MOVEMENT.T.0, positive="1")
```


##ROC Curve
```{r rf_roc}
forest_ROC_prediction <- prediction(forest_probabilities[,2], testing$MOVEMENT.T.0)
forest_ROC <- performance(forest_ROC_prediction,"tpr","fpr")
plot(forest_ROC)
```

##AUC (area under curve)
```{r rf_auc}
AUC.tmp <- performance(forest_ROC_prediction,"auc") 
forest_AUC <- as.numeric(AUC.tmp@y.values) 
forest_AUC 
```

##Lift chart
```{r rf_lift}
plotLift(forest_probabilities[,2],  testing$MOVEMENT.T.0, cumulative = TRUE, n.buckets = 10)
```

