---
title: "Project_SingRs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages
```{r, echo=FALSE}
if("pacman" %in% rownames(installed.packages()) == FALSE) {install.packages("pacman")} # Check if you have universal installer package, install if not

pacman::p_load("caret","ROCR","lift","glmnet","MASS","e1071",'dplyr',
               "partykit","rpart") #Check, and if needed install the necessary packages
library(caret)
library (ggplot2)
```

## Prediction of stock movements on stock exchange 

We will do the analysis of past share prices of the S&P500 companies, trading volumes, bid low, ask high, price volatiilty to predict the movement of the stock next day. 

Data would be used from WRDS CRSP

```{r data_loading}
data_general <- read.csv("oracle_data.csv")
data_working <- data_general
str(data_working)
```


##Clean the data
Bidlo
 - check for 0s (no price available)

Askhi
 - check for 0s (no price available)
 - check for negative values - closing ask.
 - check for positive values - highest trade

PRC
 - check for negative values - this is a bid/ask average (can ignore the negative sign)

Numtrd 
 - check for 99 (means number unavailable)
 
Create a custom function to fix missing values ("NAs") and preserve the NA info as surrogate variables

```{r, echo=TRUE}
fixNAs<-function(data_frame){
  # Define reactions to NAs
  integer_reac<-0
  factor_reac<-"FIXED_NA"
  character_reac<-"FIXED_NA"
  date_reac<-as.Date("1900-01-01")
  # Loop through columns in the data frame and depending on which class the variable is, apply the defined reaction and create a surrogate
  
  for (i in 1 : ncol(data_frame)){
    if (class(data_frame[,i]) %in% c("numeric","integer")) {
      if (any(is.na(data_frame[,i]))){
        data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
          as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
        data_frame[is.na(data_frame[,i]),i]<-integer_reac
      }
    } else
      if (class(data_frame[,i]) %in% c("factor")) {
        if (any(is.na(data_frame[,i]))){
          data_frame[,i]<-as.character(data_frame[,i])
          data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
            as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
          data_frame[is.na(data_frame[,i]),i]<-factor_reac
          data_frame[,i]<-as.factor(data_frame[,i])
          
        } 
      } else {
        if (class(data_frame[,i]) %in% c("character")) {
          if (any(is.na(data_frame[,i]))){
            data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
              as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
            data_frame[is.na(data_frame[,i]),i]<-character_reac
          }  
        } else {
          if (class(data_frame[,i]) %in% c("Date")) {
            if (any(is.na(data_frame[,i]))){
              data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
                as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
              data_frame[is.na(data_frame[,i]),i]<-date_reac
            }
          }  
        }       
      }
  } 
  return(data_frame) 
}
```
## create a custom function to search for negative values - NEEDS WORK

```{r, echo=TRUE}
fixNegatives<-function(data_frame){
  # Define reactions to Negative values
  integer_reac<-2
  factor_reac<-"NEGATIVE"
  character_reac<-"NEGATIVE"
  date_reac<-as.Date("1900-01-01")
  # Loop through columns in the data frame and depending on which class the variable is, apply the defined reaction and create a surrogate
  
  for (i in 1 : ncol(data_frame)){
    if (class(data_frame[,i]) %in% c("numeric","integer")) {
      if (any(is.numeric()<0(data_frame[,i]))){
        data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
          as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
        data_frame[is.na(data_frame[,i]),i]<-integer_reac
      }
    } else
      if (class(data_frame[,i]) %in% c("factor")) {
        if (any(is.numeric<0(data_frame[,i]))){
          data_frame[,i]<-as.character(data_frame[,i])
          data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
            as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
          data_frame[is.na(data_frame[,i]),i]<-factor_reac
          data_frame[,i]<-as.factor(data_frame[,i])
          
        } 
      } else {
        if (class(data_frame[,i]) %in% c("character")) {
          if (any(is.numeric<0(data_frame[,i]))){
            data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
              as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
            data_frame[is.na(data_frame[,i]),i]<-character_reac
          }  
        } else {
          if (class(data_frame[,i]) %in% c("Date")) {
            if (any(is.numeric<0(data_frame[,i]))){
              data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
                as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
              data_frame[is.na(data_frame[,i]),i]<-date_reac
            }
          }  
        }       
      }
  } 
  return(data_frame) 
}
```

## Create a custom function to search for 99s - NEEDS WORK
```{r, echo=TRUE}
fixNAs<-function(data_frame){
  # Define reactions to NAs
  integer_reac<-0
  factor_reac<-"FIXED_NA"
  character_reac<-"FIXED_NA"
  date_reac<-as.Date("1900-01-01")
  # Loop through columns in the data frame and depending on which class the variable is, apply the defined reaction and create a surrogate
  
  for (i in 1 : ncol(data_frame)){
    if (class(data_frame[,i]) %in% c("numeric","integer")) {
      if (any(is.integer=99(data_frame[,i]))){
        data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
          as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
        data_frame[is.na(data_frame[,i]),i]<-integer_reac
      }
    } else
      if (class(data_frame[,i]) %in% c("factor")) {
        if (any(is.integer=99(data_frame[,i]))){
          data_frame[,i]<-as.character(data_frame[,i])
          data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
            as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
          data_frame[is.na(data_frame[,i]),i]<-factor_reac
          data_frame[,i]<-as.factor(data_frame[,i])
          
        } 
      } else {
        if (class(data_frame[,i]) %in% c("character")) {
          if (any(is.integer=99(data_frame[,i]))){
            data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
              as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
            data_frame[is.na(data_frame[,i]),i]<-character_reac
          }  
        } else {
          if (class(data_frame[,i]) %in% c("Date")) {
            if (any(is.integer=99(data_frame[,i]))){
              data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
                as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
              data_frame[is.na(data_frame[,i]),i]<-date_reac
            }
          }  
        }       
      }
  } 
  return(data_frame) 
}
```

###Split the data into testing and training

```{r data_split}
data_working<-fixNAs(data_working)
data_working<-fixNegatives(data_working)

set.seed(77850) #set a random number generation seed to ensure that the split is the same everytime
inTrain <- createDataPartition(y = data_working$date,
                               p = 2000/2769, list = FALSE)
training <- data_train[ inTrain,]
testing <- data_test[ -inTrain,]
```


###Data engenieering

Transform dataset and take share price for the past 30days
Vaiarbles would be volume traded, return, volatility each day. Shareprice would be independent variable
rows - would be timeseries

