---
title: "Project_SingRs"
output: html_document
---

##Acknowledgement: Frederico Belo, Ahmed Guecioueur

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages
```{r, echo=FALSE}
if("pacman" %in% rownames(installed.packages()) == FALSE) {install.packages("pacman")} # Check if you have universal installer package, install if not

pacman::p_load("caret","ROCR","lift","glmnet","MASS","e1071",'dplyr',
               "partykit","rpart", "munsell","caTools", "caret", "randomForest") #Check, and if needed install the necessary packages
```

## Prediction of stock movements on stock exchange 

We will do the analysis of past share prices of the S&P500 companies, trading volumes, bid low, ask high, price volatiilty to predict the movement of the stock next day. 

- Data would be used from WRDS CRSP

```{r data_loading}
data_general <- read.csv("oracle_data_3.csv")
data_working <- data_general
str(data_working)
```


##Clean the data
Bidlo
 - check for 0s (no price available)
 
```{r, check Bidlo}
which(c(data_working$BIDLO.T.5,data_working$BIDLO.T.4, data_working$BIDLO.T.3, data_working$BIDLO.T.2, data_working$BIDLO.T.1, data_working$BIDLO.T.0) == 0, arr.ind = TRUE)
# confirm no 0s
```

Askhi
 - check for 0s (no price available)
```{r, check Askhi}
which(c(data_working$ASKHI.T.5,data_working$ASKHI.T.4, data_working$ASKHI.T.3, data_working$ASKHI.T.2, data_working$ASKHI.T.1, data_working$ASKHI.T.0) == 0, arr.ind = TRUE)
# Confirm no 0s
```

 
 - check for negative values - closing ask.
 - check for positive values - highest trade

PRC
 - check for negative values - this is a bid/ask average (can ignore the negative sign)

Numtrd 
 - check for 99 (means number unavailable)
```{r, check Numtrd}
which(c(data_working$NUMTRD.T.5, data_working$NUMTRD.T.4, data_working$NUMTRD.T.3, data_working$NUMTRD.T.2, data_working$NUMTRD.T.1, data_working$NUMTRD.T.0) == 99, arr.ind = TRUE)
# Confirm no 99s
```
 
Create a custom function to fix missing values ("NAs") and preserve the NA info as surrogate variables

```{r, echo=TRUE}
fixNAs<-function(data_frame){
  # Define reactions to NAs
  integer_reac<-0
  # Loop through columns in the data frame and depending on which class the variable is, apply the defined reaction and create a surrogate
  
  for (i in 1 : ncol(data_frame)){
    if (class(data_frame[,i]) %in% c("numeric","integer")) {
      if (any(is.na(data_frame[,i]))){
        data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
          as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
        data_frame[is.na(data_frame[,i]),i]<-integer_reac
      }
      }
  } 
  return(data_frame) 
}
```

## create a custom function to search for negative values - NEEDS WORK

```{r, echo=TRUE}
fixNegatives<-function(data_frame){
  # Define reactions to Negative values
  integer_reac<-2
  factor_reac<-"NEGATIVE"
  character_reac<-"NEGATIVE"
  date_reac<-as.Date("1900-01-01")
  # Loop through columns in the data frame and depending on which class the variable is, apply the defined reaction and create a surrogate
  
  for (i in 1 : ncol(data_frame)){
    if (class(data_frame[,i]) %in% c("numeric","integer")) {
      if (any(is.numeric()<0(data_frame[,i]))){
        data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
          as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
        data_frame[is.na(data_frame[,i]),i]<-integer_reac
      }
    } else
      if (class(data_frame[,i]) %in% c("factor")) {
        if (any(is.numeric<0(data_frame[,i]))){
          data_frame[,i]<-as.character(data_frame[,i])
          data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
            as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
          data_frame[is.na(data_frame[,i]),i]<-factor_reac
          data_frame[,i]<-as.factor(data_frame[,i])
          
        } 
      } else {
        if (class(data_frame[,i]) %in% c("character")) {
          if (any(is.numeric<0(data_frame[,i]))){
            data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
              as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
            data_frame[is.na(data_frame[,i]),i]<-character_reac
          }  
        } else {
          if (class(data_frame[,i]) %in% c("Date")) {
            if (any(is.numeric<0(data_frame[,i]))){
              data_frame[,paste0(colnames(data_frame)[i],"_surrogate")]<-
                as.factor(ifelse(is.na(data_frame[,i]),"1","0"))
              data_frame[is.na(data_frame[,i]),i]<-date_reac
            }
          }  
        }       
      }
  } 
  return(data_frame) 
}
```

###Split the data into testing and training
```{r data_split_cleaning}
#data_working$date <- NULL

data_working$date <- as.Date(data_working$date, "%d-%b-%y")
data_working$VOL.T.10 <- as.numeric(data_working$VOL.T.10)
data_working$VOL.T.9 <- as.numeric(data_working$VOL.T.9)
data_working$VOL.T.8 <- as.numeric(data_working$VOL.T.8)
data_working$VOL.T.7 <- as.numeric(data_working$VOL.T.7)
data_working$VOL.T.6 <- as.numeric(data_working$VOL.T.6)
data_working$VOL.T.5 <- as.numeric(data_working$VOL.T.5)
data_working$VOL.T.4 <- as.numeric(data_working$VOL.T.4)
data_working$VOL.T.3 <- as.numeric(data_working$VOL.T.3)
data_working$VOL.T.2 <- as.numeric(data_working$VOL.T.2)
data_working$VOL.T.1 <- as.numeric(data_working$VOL.T.1)
data_working$MOVEMENT.T.0 <- as.factor(data_working$MOVEMENT.T.0)

#data_working_wT0 <- data_working[,1:50]
#data_working_wT0 <- cbind(data_working_wT0, data_working$MOVEMENT.T.0)
#names(data_working_wT0)[51] <- paste("MOVEMENT.T.0")

train_general <- data_working[1:2000,]
test_general <- data_working[2001:2759,]
test_general_womov <- test_general
test_general_womov$MOVEMENT.T.0 <- NULL
```


```{r data_split_2}
#data_working<-fixNAs(data_working)
#data_working<-fixNegatives(data_working)

set.seed(77850) #set a random number generation seed to ensure that the split is the same everytime
inTrain <- createDataPartition(y = train_general$MOVEMENT.T.0,
                               p = 0.75, list = FALSE)
training <- train_general[ inTrain,]
testing <- train_general[ -inTrain,]
```


###Data engenieering

Transform dataset and take share price for the past 30days
Vaiarbles would be volume traded, return, volatility each day. Shareprice would be independent variable
rows - would be timeseries

###Random forest


```{r random_forest}
start_time <- Sys.time()
model_forest <- randomForest(MOVEMENT.T.0~ ., data=training, importance=TRUE, proximity=TRUE, type="classification")
print(model_forest)
end_time <- Sys.time()
end_time - start_time

plot(model_forest)
importance(model_forest)
varImpPlot(model_forest)
```


Finding predicitons: probabilities and classification


```{r rf_predictions}
forest_probabilities<-predict(model_forest,newdata=testing,type="prob") 
forest_classification<-rep("0",length(testing$MOVEMENT.T.0))
forest_classification[forest_probabilities[,2]>0.5]="1"
forest_classification<-as.factor(forest_classification)
```

## RF Confusion matrix
Display confusion matrix. Note, confusion matrix actually displays a better accuracy with threshold of 50%
```{r rf_cf}
rf_cm<-confusionMatrix(forest_classification,testing$MOVEMENT.T.0, positive="1")
rf_cm
```

##ROC Curve
```{r rf_roc}
forest_ROC_prediction <- prediction(forest_probabilities[,2], testing$MOVEMENT.T.0)
forest_ROC <- performance(forest_ROC_prediction,"tpr","fpr")
plot(forest_ROC)
```

##AUC (area under curve)
```{r rf_auc}
AUC.tmp <- performance(forest_ROC_prediction,"auc") 
forest_AUC <- as.numeric(AUC.tmp@y.values) 
forest_AUC 
```

##Lift chart
```{r rf_lift}
plotLift(forest_probabilities[,2],  testing$MOVEMENT.T.0, cumulative = TRUE, n.buckets = 10)
```

