---
title: "Project_SingRs"
output: html_document
---

##Acknowledgement: Frederico Belo, Ahmed Guecioueur
###Reproduction 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages
```{r, echo=FALSE}
if("pacman" %in% rownames(installed.packages()) == FALSE) {install.packages("pacman")} # Check if you have universal installer package, install if not

pacman::p_load("caret","ROCR","lift","glmnet","MASS","e1071",'dplyr',
               "partykit","rpart", "munsell","caTools", "caret", "randomForest") #Check, and if needed install the necessary packages
```

## Prediction of stock Movements on stock exchange 

We will do the analysis of past share prices of the S&P500 companies, trading volumes, bid low, ask high, price volatiilty to predict the Movement of the stock next day. 

- Data would be used from WRDS CRSP

```{r data_loading}
data_general <- read.csv("vixcurrent.csv")
data_working <- data_general
str(data_working)
```


###Split the data into testing and training
```{r data_split_cleaning}
data_working$Date <- as.Date(as.character(data_working$Date), "%d/%m/%y")
data_working <- ts(data_working, start="01-02-2004", frequency = 252)
```


```{r time_series_analysis}
vix_model <- ets(data_working, model = "ZZZ")




```

```{r hmm}
library(ldhmm)
library(moments)
ts = ldhmm.ts_log_rtn("spx", on ="days")
sapply(0:10, function (drop) kurtosis(ldhmm.drop_outliers(ts$x, drop)))
ldhmm.ts_abs_acf(ts$x,drop =0 , lag.max =6)
ldhmm.ts_abs_acf(ts$x,drop =10, lag.max =6)


mu_1 = 0.0006
sigma_1 = 0.01
mu_2 = −0.0007
sigma_2 = 0.02
m = 2

param0=matrix(c(mu_1, sigma_1, mu_2, sigma_2), m, 2, byrow=TRUE)
gamma0=ldhmm.gamma_init(m)
h <- ldhmm(m, param0, gamma0, stationary=TRUE)
hd <- ldhmm.mle(h, ts$x, decode = TRUE, print.level =2)
hd@param
hd@gamma
hd@delta

ldhmm.ld_stats(hd)

hd@states.local.stats

ldhmm.ld_stats(hd)

ldhmm.calc_stats_from_obs(hd, drop=11)

ldhmm.oxford_man_plot_obs(hd)

```



```{r hmm_2}
states_prob <- t(as.data.frame(hd@states.global.stats))

write.csv(states_prob,"states_prob.csv")


```

```{r hmm_3}

hs<-ldhmm.simulate_state_transition(hd, init =100000)
kurtosis(hs@observations)

ldhmm.simulate_abs_acf(hd, n=100000, lag.max =1)

```
